Capstone Movielens Report

### 2025-12-13 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `calc_UMGY_SmoothedDay_effect.MSE.cv`.

::: { #func.calc_UMGY_effect_MSE.cv.args.UMGY_effect .sidebar }
  - **UMGY_effect:** Data frame representing the *UMGY Effect Model*.

@@

Validates the *UMGYD Effect (UMGYDE) Model* using data passed in the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect) argument on the *test dataset* passed in the [test_set](#func.calc_UMGY_SmoothedDay_effect.MSE.args.test_set) argument.

@@

In *this Project*, the *UMGYDE* data object representing the *UMGYDE Model* (and being passed in the [day_smoothed_effect](#func.loess_de.args.de_bias.dat) argument) is created and returned by the following functions described above in *this Section*:
  - [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) 
  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect) 
  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

Internally, this function calls the auxiliary functions [UMGY_SmoothedDay_effect.predict](#func.UMGY_SmoothedDay_effect.predict) (described above in this section), passing to it the value of the argument [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect) in the same-named argument.

@@

::: {.noteblock data-latex=""}
This helper function is used internally in the [calc_UMGY_effect_RMSE.cv](#func.calc_UMGY_effect_RMSE.cv) function (described below in this section) as shown in the [Usage](#func.calc_UMGY_SmoothedDay_effect.MSE.cv.usage) subsection of *This Function's Description Section* below.
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_UMGY_SmoothedDay_effect.MSE.cv` is called internally from the [calc_UMGY_effect_RMSE.cv](#func.calc_UMGY_effect_RMSE.cv) function described below in this section.
:::
:::

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.cv.args.UMGYD_effect .sidebar }
  - **UMGYD_effect:** Data frame representing the *UMGYDE Model*.

@@

The function calculates *MSE* for the *UMGYDE Model* passed in the [UMGYD_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.cv.args.UMGYD_effect) argument using the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

::: {.noteblock data-latex=""}
This function internally calls the [calc_UMGY_SmoothedDay_effect.MSE](#func.calc_UMGY_SmoothedDay_effect.MSE) function for each iteration of the *K-Fold Cross-Validation*, calculating the *MSE* for the corresponding validation dataset of the [`edx_CV` Object], and then calculates the mean of the obtained results.

@@

The simplified version of the source code of the [calc_UMGY_SmoothedDay_effect.MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L147) function is shown below:
```{r eval=FALSE}
```

>>>>

::: { #func.calc_UMGY_effect_MSE.cv.args.UMGY_effect .sidebar }
  - **UMGY_effect:** Dataframe object representing the *UMGY Effect Model*.

@@

Validates the *UMGYD Effect (UMGYDE) Model* represented by the data passed in the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect) argument on the *test dataset* passed in the [test_set](#func.calc_UMGY_SmoothedDay_effect.MSE.args.test_set) argument.

@@

In *this Project*, the *UMGYDE* data object representing the *UMGYDE Model* (and being passed in the [day_smoothed_effect](#func.loess_de.args.de_bias.dat) argument) is created and returned by the following functions described above in *this Section*:

  - [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect)

  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)

Internally, this function calls the auxiliary functions [UMGY_SmoothedDay_effect.predict](#func.UMGY_SmoothedDay_effect.predict) (described above in this section), passing to it the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect) value in the same-named argument.

@@

Validates the *UMGYD Effect (UMGYDE) Model* represented by the data object passed in the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.cv.args.day_smoothed_effect) argument with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV] (we use $K = 5$ in *this Project*).

@@

```{r eval=FALSE}
  sqrt(calc_UMGY_SmoothedDay_effect.MSE.cv(day_smoothed_effect))
```

::: {.noteblock data-latex=""}
In the code snippet above, the function `calc_UMGY_SmoothedDay_effect.MSE.cv` is called internally from the [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) function described below in this section.
:::
:::

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.cv.args.day_smoothed_effect .sidebar }
  - **day_smoothed_effect:** Dataframe object representing the *UMGYDE Model*.

@@

This function commits the *UMGYDE Model* validation by calculating the *Mean Squared Error (MSE)* on the validation datasets contained in the [`edx_CV` Object] described below in [Appendix B: Models Training Datasets].

::: {.noteblock data-latex=""}
This function internally calls the [calc_UMGY_SmoothedDay_effect.MSE](#func.calc_UMGY_SmoothedDay_effect.MSE) function for each iteration of the *K-Fold Cross-Validation*, passing [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.cv.args.day_smoothed_effect) value in the same-named argument to the called function and calculating the *MSE* for the corresponding validation dataset of the [`edx_CV` Object]. After that, it calculates the mean of the obtained results.

The $K$ value for the *K-Fold Cross-Validation* is determined by the length of the [`edx_CV` Object] (in *this Project* we use $K = 5$).
:::

In *this Project*, the *UMGYDE* data object representing the *UMGYDE Model* (and being passed in the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.cv.args.day_smoothed_effect) argument) is created and returned by the following functions described above in *this Section*:

  - [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect)

  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)

::: {.noteblock data-latex=""}
This is an auxiliary function that is called internally from the [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) function (described below in this section) as shown in the [Usage](#func.calc_UMGY_SmoothedDay_effect.MSE.usage) section of *This Description* above.
:::
:::

@@

The slightly simplified version of the source code of the [calc_UMGY_SmoothedDay_effect.MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L147) function is shown below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.MSE.cv <- function(day_smoothed_effect){
  smth_day_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> 
      calc_UMGY_SmoothedDay_effect.MSE(day_smoothed_effect)
  })
  put_end_date(start)
  put_log1("Function `calc_UMGY_SmoothedDay_effect.MSE.cv`:
MSE value have been computed for the %1-Fold Cross Validation samples.", 
           CVFolds_N)
  mean(smth_day_effect_MSEs)
}
```
